---
title: "Assignment 2: Oregon_Registration_Create_Tidyfile"
author: "Paul Gronke"
date: "February 10, 2017"
output: html_document
---

This file reads in the five Oregon voter registration files, binds them together, then selects
only key data elements, and saves a much smaller file for class use.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(car)      # John Fox program for his regression text, nice recode command
library(rms)      # Frank Harrell regression modeling strategies package
library(Hmisc)    # Frank Harrell miscellaneous commands package
library(broom)    # Create "tidy" data frames from statistical results
library(popbio)   # Matrix population package, has beautiful logit+histogram plot command

```

## Read and bind the files

Read the files from Google Drive, bind them together

```{r file_read, cache = TRUE}

## This directory is on Paul's computer

setwd("~/Google Drive/Teaching/Bray_Gronke_Data_Science/Oregon Voter Files/January 2017/StatewideVoterList_January2017/VoterList_Jan2017")
filenames <- list.files(pattern = "2017.txt")
files <- lapply(filenames, read_tsv, col_names=TRUE, 
                col_types=paste(rep("c",40),collapse=""),
                trim_ws=TRUE)

or_voter <- bind_rows(files)

## Make this our primary data frame
or_voter <- as.data.frame(or_voter)


```

## Trimming the file

Remove extraneous data elements

```{r trim , cache = TRUE}
or_voter$CONFIDENTIAL <- as.factor(or_voter$CONFIDENTIAL)
or_voter$COUNTY <- as.factor(or_voter$COUNTY)
# or_voter$ZIP_CODE <- as.factor(or_voter$ZIP_CODE)
or_voter$PARTY_CODE <- as.factor(or_voter$PARTY_CODE)
or_voter$STATUS <- as.factor(or_voter$STATUS)

str(or_voter)

or_reg_tidyfile <- or_voter %>% select(VOTER_ID, BIRTH_DATE, CONFIDENTIAL, EFF_REGN_DATE, STATUS, PARTY_CODE, COUNTY, ZIP_CODE,PRECINCT_NAME, PRECINCT, SPLIT) %>%
    filter(VOTER_ID !="ACP")
str(or_reg_tidyfile)

write_csv(or_reg_tidyfile, "~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/or_reg_tidyfile.csv")
rm(or_reg)

```

## Read in the Oregon OMV file

Read in the data file that contains information on individuals who were registered automatically

```{r omv_file , cache = TRUE}

setwd("~/Google Drive/Teaching/Bray_Gronke_Data_Science/Oregon Voter Files/January 2017/StatewideVoterList_January2017")

#
# Couldn't figure out how to deal with this using read_table, it kept choking on the quotations
#

omv_jan2017 <- read.table("OMV_Jan2017.txt", header = TRUE, skip = 0)
str(omv_jan2017)
omv_tidyfile <- omv_jan2017 %>% select(VOTER_ID, DESCRIPTION, COUNTY) %>% filter(VOTER_ID != "ACP")
str(omv_tidyfile)

write_csv(omv_tidyfile, "~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/omv_tidyfile.csv")
rm(or_omv)

```

## Read in the Oregon Voter History file

```{r history_file, cache = TRUE}

setwd("~/Google Drive/Teaching/Bray_Gronke_Data_Science/Oregon Voter Files/January 2017/StatewideVoterList_January2017/VoterHistory_Jan2017/")

filenames <- list.files(pattern = "2017.txt")
files <- lapply(filenames,read_delim, 
                col_types = paste(rep("c", 65), collapse = ""),
                delim = "\t", 
                escape_double = FALSE, 
                trim_ws = TRUE)

or_history <- bind_rows(files)  %>%
    select(VOTER_ID, BIRTH_DATE, EFF_REGN_DATE, 41:65) 
    
write_csv(or_history, "~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/votehistory_tidyfile.csv")
rm(files)
rm(or_history)

```

## Join Step

```{r join_step, cache = TRUE}

or_reg1 <- read_csv("~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/or_reg_tidyfile.csv")

or_omv1 <- read_csv("~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/omv_tidyfile.csv")

or_history1 <- read_csv("~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/votehistory_tidyfile.csv")

or_voter <- or_reg1 %>% 
  distinct(VOTER_ID, .keep_all = TRUE) %>% 
  left_join(
    distinct(or_omv1, VOTER_ID, .keep_all = TRUE),
    by = "VOTER_ID")

or_voter <- or_voter %>% 
  left_join(
    distinct(or_history1, VOTER_ID, .keep_all = TRUE),
    by = "VOTER_ID")

write_csv(or_voter, "~/Google Drive/Teaching/Bray_Gronke_Data_Science/gronke_data_files/or_voter.csv")
```

## Recodes

```{r logit, cache = TRUE}

# Create birth date and age variable, remove bad birth years

or_voter <- or_voter %>% mutate(BIRTH_DATE = mdy(BIRTH_DATE.x)) %>% filter(BIRTH_DATE > as_date("1902-1-1"))
or_voter <- or_voter %>% mutate(age = 2017-year(BIRTH_DATE))
or_voter <- or_voter %>% mutate(age2014 = 2014-year(BIRTH_DATE))

# Create vote variable November 2016
table(or_voter$`11/08/2016`)
or_voter$vote_nov2016 <- ifelse(or_voter$`11/08/2016` == "YES", 1, ifelse(or_voter$`11/08/2016` == "NO", 0, NA))
table(or_voter$vote_nov2016)

#Create vote variable for November 2014 
table(or_voter$`11/04/2014`)
or_voter$vote_nov2014 <- ifelse(or_voter$`11/04/2014` == "YES", 1, ifelse(or_voter$`11/04/2014` == "NO", 0, NA))
table(or_voter$vote_nov2014)

# Need to have Fox's car package installed to make this command work
or_voter$newparty1 <- as.factor(recode(or_voter$PARTY_CODE,"'DEM' = 'Democrat'; 'REP' = 'Republican'; 
                      'NAV' = 'Non-Affiliated' ; else = 'Other'"))
table(or_voter$newparty1)

```

## Logit Models 2016 election

```{r rms, cache = TRUE}

# Base R results

m1 <- glm(vote_nov2016 ~ age, data = or_voter, family = binomial)
m2 <- glm(vote_nov2016 ~ age + newparty1, data = or_voter, family = binomial)
summary(m1)
coef(m1)

# Prettier results using Broom

tidy(m1)
glance(m1)

tidy(m2)
glance(m2)

# Prettier results using rms logistic regression model command

m3 <- lrm(vote_nov2016 ~ age, data=or_voter)
m3

m4 <- lrm(vote_nov2016 ~ age + newparty1, data=or_voter, penalty=5)
m4

```

## Logit Models: Examining Effects Using John Fox "effects" package

```{r logit2}

library(effects)

# This allows for easy calculation of marginal effects
m1.all <- allEffects(m1)
print(m1.all, digits=3)
plot(m1.all)

m2.all <- allEffects(m2)
print(m2.all, digits=3)
plot(m2.all)
```

## Make some prediction tables, there must be a better way to do this 

```{r predict}

# Prediction table for party
predtable <- with(or_voter, data.frame(age=mean(age), newparty1=factor("Democrat", 
                    "Non-Affiliated", "Republican", "Other")))
predtable$voteP <- predict(m2, newdata=predtable, type = "response")                 
predtable

# Prediction table for 18 year olds, by party
predtable2 <- with(or_voter, data.frame(age=18, newparty1=factor("Democrat", 
                    "Non-Affiliated", "Republican", "Other")))
predtable2$voteP <- predict(m2, newdata=predtable2, type = "response")                 
predtable2

#Prediction table for 65 year olds, by party
predtable3 <- with(or_voter, data.frame(age=65, newparty1=factor("Democrat", 
                    "Non-Affiliated", "Republican", "Other")))
predtable3$voteP <- predict(m2, newdata=predtable3, type = "response")                 
predtable3

```


## Pretty Plot Time 

```{r plots}

# ###############
# Start with base R plot. This is really slow 
# #################

plot(or_voter$age, or_voter$vote_nov2016, xlab="Age", ylab = "Voted Nov 2016", 
      xlim=c(18,100), ylim=c(-0.1, 1.1), mgp=c(2, .5, 0))
curve(exp(coef(m1)[1] + coef(m1)[2]*x)/(1 + exp(coef(m1)[1] + coef(m1)[2]*x)), add=TRUE)

# ###################################
# Now plot predicted values directly
# ##################################

  # Start by setting a range for our predictions
newdata2 <- with(or_voter, data.frame(age = 18:90))
  # Put predicted values into this new data frame
preds <- predict(m1, newdata2, type="response", se.fit = TRUE)
  # Vector of fitted values
predf <- preds$fit
  # Vector of predicted values - 1.96 SE
lower <- preds$fit - (1.96*preds$se.fit)
  # Vector of predicted values + 1.96 SE
upper <- preds$fit + (1.96*preds$se.fit)
  # Now plot it, start with the plot, then add SE bounds
plot(18:90, predf, type="l", ylab = "Predicted Probabilty of Voting", xlab = "Age", bty = "n")
lines(18:90, lower, lty=2)
lines(18:90, upper, lty=2)

# PopBio Package has a nice canned version of a cool probabilty plot + histogram of values, a
#   very informative plot

logi.hist.plot(or_voter$age, or_voter$vote_nov2016, boxp=FALSE, type="hist", col="gray")
```

## Voter turnout in 2014 Midterm election

```{r midterm, cache = TRUE}

m5 <- lrm(vote_nov2014 ~ age, data=or_voter)
m5

m6 <- lrm(vote_nov2014 ~ age + newparty1, data=or_voter, penalty=5)
m6

```
