title: "OMV file 1"
author: "RosaKalish"
output: github_document
========================================================

Load Packages
```{r}
library(tidyverse)
library(lubridate)
library(car)      # John Fox program for his regression text, nice recode command
library(rms)      # Frank Harrell regression modeling strategies package
library(Hmisc)    # Frank Harrell miscellaneous commands package
library(broom)    # Create "tidy" data frames from statistical results
library(popbio)   # Matrix population package, has beautiful logit+histogram plot command
library(tidyverse)
library(data.table)
library(stringr)
library(tidyr)

```{r}
```
Import and Bind Voter Data 
```{r}

#vote_particip <- read_csv("http://bit.ly/2kG37yJ")
vote_motor <- read_csv("http://bit.ly/2lCadlB")


#These are the files with correct voting turnout data, however they are still missing registration type, which I join from the data used in Home Work 1. 
file.choose()
voter1 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD1_VoterHistory_Jan2017.txt")
voter2 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD2_VoterHistory_Jan2017.txt")
voter3 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD3_VoterHistory_Jan2017.txt")
voter4 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD4_VoterHistory_Jan2017.txt")
voter5 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD5_VoterHistory_Jan2017.txt")

voter_all <- rbind(voter1, voter2, voter3, voter4, voter5)
---------
#These are the voter files I downloaded from Pauls google drive, I am not currently using them. 
  
#voter31 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD1_VoterList_Jan2017.txt")
#voter32 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD2_VoterList_Jan2017.txt")
#voter33 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD3_VoterList_Jan2017.txt")
#voter34 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD4_VoterList_Jan2017.txt")
#voter35 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD5_VoterList_Jan2017.txt")

#voter3_all <- rbind(voter1, voter2, voter3, voter4, voter5)
 

----
#voter_all <- inner_join(x = voter2_all,y = vote_motor, by = "VOTER_ID")

voter2_all <- left_join(voter_all, vote_motor, by = "VOTER_ID")
```
Tidy Vote File 
```{r}

voter_or <- voter2_all %>%
  select(VOTER_ID, FIRST_NAME, LAST_NAME, COUNTY.x, CITY, BIRTH_DATE, STATUS, PARTY_CODE, PRECINCT_NAME, PRECINCT, ZIP_CODE, `11/08/2016`, DESCRIPTION, `11/06/2012`, `11/04/2008`)

```
Import Geographic Data
```{r}
#I first import Census tract data, then State Legislative district, then County. I finally import Zip Code which I use for this project. However, while the others are from the 2015 community report, the Zip Code data comes from the 2010 election, so there is not perfect demographic information. 
#file.choose()
#census <- read_csv("/Users/rosa/Desktop/SEcensus.csv")

#file.choose()
#stateleg <- read_csv("/Users/rosa/Desktop/SEstateleg.csv")

#file.choose()
#county <- read_csv("/Users/rosa/Desktop/SEcounty.csv")

#file.choose()
zipcode <- read_csv("/Users/rosa/Desktop/zipcode.csv")
```
Tidy Geographic Data 
```{r}

zipcode <- zipcode %>%
  mutate(Geo_NAME = str_replace_all(Geo_NAME, pattern = "ZCTA5 ", replacement = ""))

# Make note of which zips have "(part)" - For later
part_zips <- zipcode %>%
  filter(grepl("part", Geo_NAME))

# Clear those out
zipcode <- zipcode %>%
    mutate(Geo_NAME = substr(Geo_NAME, 1, 5)) %>%
  rename(total_pop = SE_T003_001,
         male = SE_T003_002,
         female = SE_T003_003,
         total_pop2 = SE_T054_001,
         white = SE_T054_002,
         black = SE_T054_003, 
         ai = SE_T054_004,
         asian = SE_T054_005,
         hawaiian = SE_T054_006,
         other = SE_T054_007, 
         two = SE_T054_008,
         ZIP_CODE = Geo_NAME) 

         
# Just Sex
sex <- zipcode %>%
  select(ZIP_CODE, total_pop, male, female)

# Just Race

zipcode$p_black <- zipcode$black / zipcode$total_pop
zipcode$p_white <- zipcode$white / zipcode$total_pop

race <- zipcode %>%
  select(ZIP_CODE, total_pop, white, black, ai, asian, hawaiian, other, two, p_black, p_white)

```

Get at zip code level for Oregon registered voters
```{r}
total_regs <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(count = n())
```
Construct Variables for Analysis: 

Proportion of registered that voted on Nov 2016 and Nov 2012 
```{r}
prop_voted2016 <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_v16 = mean(`11/08/2016` == "YES"))

prop_voted2012 <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_v12 = mean(`11/06/2012` == "YES"))
```
Proportion that are not Motor Voter Registered 
```{r}
a <- (is.na(voter_or$DESCRIPTION))
voter_or$DESCRIPTION[a] <- "conventional"
voter_or$DESCRIPTION <- as.factor(voter_or$DESCRIPTION)

table(voter_or$DESCRIPTION)

prop_regcon <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_reg = mean(DESCRIPTION == "conventional"))

#Dont need this any more, also does not work. 
#votor_or2 <- votor_or %>%
#  mutate(DESCRIPTION = stringr::str_replace_all(DESCRIPTION, NA, "Not Registered")
         
#MV <- voter_or2 %>%
  #group_by(ZIP_CODE) %>%
  #summarize(prop_MV = mean(DESCRIPTION != "conventional"))
  
  
      # prop_motorvoter <- voter_or %>%
 # group_by(ZIP_CODE) %>%
 # summarize(prop_NotMV = mean(`DESCRIPTION` == "Not Registered"))
```

Construct data sets for analysis with Voter info by Zip Code including, number of registered voters, how they registered, and the proportion that voted.
```{r}
zipcode_data <- inner_join(x = total_regs,
                           y = prop_voted2016,
                           by = "ZIP_CODE")

zipcode_data <- inner_join(x = zipcode_data,
                           y = prop_voted2012,
                           by = "ZIP_CODE")

zipcode_data <- inner_join(x = zipcode_data,
                           y = prop_regcon,
                           by = "ZIP_CODE")

# Join sex with voter reg aggregated data
sex_reg <- inner_join(x = sex, y = zipcode_data, 
                      by = "ZIP_CODE")

# Join race with voter reg aggregated data
race_reg <- inner_join(x = race, y = zipcode_data, 
                      by = "ZIP_CODE")
# One set with both 

zip_reg <- inner_join(x = race_reg, y = sex, 
                      by = "ZIP_CODE")

# Add Turnout difference 
require(dplyr)
require(tidyr)
mydata = tbl_df(mydata)

zip_reg <- zip_reg %>%
mutate (vote_diff = prop_v16 / prop_v12)
```
Modeling: 

Model 1 and 2. Race and Registration
```{r}
# Modle 1 and 2: Race and Registration Type 
 ggplot(zip_reg, aes(x = p_black, y = prop_reg)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "mediumpurple2") +
    ggtitle("Registration Type by Zip Code Blackness") +
  labs(x="Percet Black of Zip Code",y="Proportion of Voters Registered Traditionally")
  
 ggplot(zip_reg, aes(x = p_white, y = prop_reg)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "mediumpurple2") +
    ggtitle("Registration Type by Zip Code Whiteness") +
  labs(x="Percet White of Zip Code",y="Proportion of Voters Registered Traditionally")
  
m1 <- lm(p_white ~ propb_reg, data = zip_reg)
summary(m1)

m2 <- lm(p_black ~ prob_reg, data = zip_reg)
summary(m2)

#this suggests that the probability of being registered to vote not through OMV, is higher for counties with both higher black and white populations. This may be explained by the fact that there are such small percentages of black voters in oregon. 

```
Models Continued 
```{r}
# Model 3: Registration type and Voter turnout 
 ggplot(zip_reg, aes(x = prop_reg, y = vote_diff)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "palegreen") +
    ggtitle("Voting Increase and Registration Type across Zip Code") +
  labs(x="Proportion of Voters Registered Traditionally",y="Increase in Turnout for 2016 Election")+
   ylim(0, 3)

m3 <-lm(vote_diff ~ prop_reg, data = zip_reg)
summary(m3)
#Why wont this model work? 

# Model 4: Population and Registration Type 
 ggplot(zip_reg, aes(x = total_pop.x, y = prop_reg)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "orange") +
    ggtitle("AVR Prevalance across ZipCode Populaiton") +
  labs(x="ZipCode Population ",y="Proportion of Voters Registered Traditionally")

m4 <-lm(total_pop.x ~ prop_reg, data = zip_reg)
summary(m4)


# Model 5: population and turnout rate
ggplot(zip_reg, aes(x = total_pop.x, y = vote_diff)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "yellowgreen") +
    ggtitle("Turnout Rates by ZipCode Populaiton") +
  labs(x="ZipCode Population ",y="Change in Turnout Rates from 2012 and 2016 Presidential Election")+
   ylim(0, 2)

m5 <-lm(total_pop.x ~ vote_diff, data = zip_reg)
summary(m5)

# Model 6 and 7: turnout by race 
ggplot(zip_reg, aes(x = p_white, y = vote_diff)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "skyblue1") +
    ggtitle("Turnout Rates by Zipcode Whitness") +
  labs(x="ZipCode Whiteness ",y="Change in Turnout Rates from 2012 and 2016 Presidential Election")+
   ylim(0, 2)

ggplot(zip_reg, aes(x = p_black, y = vote_diff)) +
geom_point(color= "hotpink")+
  geom_smooth(method=`lm`, color= "skyblue1") +
    ggtitle("Turnout Rates by Zipcode Blackness") +
  labs(x="ZipCode Blackness ",y="Change in Turnout Rates from 2012 and 2016 Presidential Election")+
   ylim(0, 2)

m6 <- lm(vote_diff ~ p_white, data = zip_reg)
summary(m6)


m7 <- lm(vote_diff ~ p_black, data = zip_reg)
summary(m7)

```
Extra Code 
```{r}
#race_reg2<- race_reg %>% 
  #gather(key="race", value = "number", 3:9)

#sex_reg2 <- sex_reg %>% 
 # gather(key="sex", value = "number", 3:4)

#both_regtidy <-inner_join(x = race_reg2, y = sex_reg2, 
                    #  by = "ZIP_CODE") 


```

