title: "OMV file 1"
author: "RosaKalish"
output: github_document
========================================================

Load Packages
```{r}
library(tidyverse)
library(lubridate)
library(car)      # John Fox program for his regression text, nice recode command
library(rms)      # Frank Harrell regression modeling strategies package
library(Hmisc)    # Frank Harrell miscellaneous commands package
library(broom)    # Create "tidy" data frames from statistical results
library(popbio)   # Matrix population package, has beautiful logit+histogram plot command

```{r}
```
Import and Bind 
```{r}

#file.choose()
voter1 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD1_VoterList_Jan2017.txt")
voter2 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD2_VoterList_Jan2017.txt")
voter3 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD3_VoterList_Jan2017.txt")
voter4 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD4_VoterList_Jan2017.txt")
voter5 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD5_VoterList_Jan2017.txt")

voter_all <- rbind(voter1, voter2, voter3, voter4, voter5)
head(voter_all) 

#file.choose()
spatial <- read_csv("/Users/rosa/Downloads/SPATIAL_INFO.csv")
```
Tidy Vote FIle 
```{r}
voter_or <- voter_all %>%
  select(VOTER_ID, FIRST_NAME, LAST_NAME, COUNTY, CITY, BIRTH_DATE, STATUS, PARTY_CODE, RES_ADDRESS_1, PRECINCT_NAME, PRECINCT, ZIP_CODE) 
```
Import Geographic dta 
```{r}
#file.choose()
geo <- read_csv("/Users/rosa/Desktop/socialexplorer-zipcode.csv")
voter_geo <- inner_join(voter_or, spatial) 
by COUNTY
```

```{r}
voter_or2 <- voter_or %>%
  mutate(COUNTY = tolower(COUNTY))
spatial2 <- spatial %>%
  mutate(COUNTY = tolower(COUNTY))
join <- inner_join(voter_or2, spatial2, by = "COUNTY")
```

```{r}
voter_or %>% count(COUNTY)
voter_or %>% count(COUNTY, PARTY_CODE)
```
What info is in the voter_all and voter_or files, what should I be looking at and how should I tidy? 
