title: "OMV file 1"
author: "RosaKalish"
output: github_document
========================================================

Load Packages
```{r}
library(tidyverse)
library(lubridate)
library(car)      # John Fox program for his regression text, nice recode command
library(rms)      # Frank Harrell regression modeling strategies package
library(Hmisc)    # Frank Harrell miscellaneous commands package
library(broom)    # Create "tidy" data frames from statistical results
library(popbio)   # Matrix population package, has beautiful logit+histogram plot command
library(tidyverse)
library(data.table)
library(stringr)
library(tidyr)

```{r}
```
Import and Bind Voter Data 
```{r}

#vote_particip <- read_csv("http://bit.ly/2kG37yJ")
vote_motor <- read_csv("http://bit.ly/2lCadlB")


#These are the files with correct voting turnout data, however they are still missing registration type, which I join from the data used in Home Work 1. 
file.choose()
voter1 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD1_VoterHistory_Jan2017.txt")
voter2 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD2_VoterHistory_Jan2017.txt")
voter3 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD3_VoterHistory_Jan2017.txt")
voter4 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD4_VoterHistory_Jan2017.txt")
voter5 <- read_tsv("/Users/rosa/Desktop/or_voter_history/CD5_VoterHistory_Jan2017.txt")

voter_all <- rbind(voter1, voter2, voter3, voter4, voter5)
---------
#These are the voter files I downloaded from Pauls google drive, I am not currently using them. 
  
#voter31 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD1_VoterList_Jan2017.txt")
#voter32 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD2_VoterList_Jan2017.txt")
#voter33 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD3_VoterList_Jan2017.txt")
#voter34 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD4_VoterList_Jan2017.txt")
#voter35 <- read_tsv("/Users/rosa/Documents/OMV project/data /CD5_VoterList_Jan2017.txt")

#voter3_all <- rbind(voter1, voter2, voter3, voter4, voter5)
 

----
#voter_all <- inner_join(x = voter2_all,y = vote_motor, by = "VOTER_ID")

voter2_all <- left_join(voter_all, vote_motor, by = "VOTER_ID")
```
Tidy Vote File 
```{r}

voter_or <- voter2_all %>%
  select(VOTER_ID, FIRST_NAME, LAST_NAME, COUNTY.x, CITY, BIRTH_DATE, STATUS, PARTY_CODE, PRECINCT_NAME, PRECINCT, ZIP_CODE, `11/08/2016`, DESCRIPTION, `11/06/2012`, `11/04/2008`)

```
Import Geographic Data
```{r}
#I first import Census tract data, then State Legislative district, then County. I finally import Zip Code which I use for this project. However, while the others are from the 2015 community report, the Zip Code data comes from the 2010 election, so there is not perfect demographic information. 
#file.choose()
#census <- read_csv("/Users/rosa/Desktop/SEcensus.csv")

#file.choose()
#stateleg <- read_csv("/Users/rosa/Desktop/SEstateleg.csv")

#file.choose()
#county <- read_csv("/Users/rosa/Desktop/SEcounty.csv")

#file.choose()
zipcode <- read_csv("/Users/rosa/Desktop/zipcode.csv")
```
Tidy Geographic Data 
```{r}

zipcode <- zipcode %>%
  mutate(Geo_NAME = str_replace_all(Geo_NAME, pattern = "ZCTA5 ", replacement = ""))

# Make note of which zips have "(part)" - For later
part_zips <- zipcode %>%
  filter(grepl("part", Geo_NAME))

# Clear those out
zipcode <- zipcode %>%
    mutate(Geo_NAME = substr(Geo_NAME, 1, 5)) %>%
  rename(total_pop = SE_T003_001,
         male = SE_T003_002,
         female = SE_T003_003,
         total_pop2 = SE_T054_001,
         white = SE_T054_002,
         black = SE_T054_003, 
         ai = SE_T054_004,
         asian = SE_T054_005,
         hawaiian = SE_T054_006,
         other = SE_T054_007, 
         two = SE_T054_008,
         ZIP_CODE = Geo_NAME) 

         
# Just Sex
sex <- zipcode %>%
  select(ZIP_CODE, total_pop, male, female)

# Just Race

zipcode$p_black <- zipcode$black / zipcode$total_pop
zipcode$p_white <- zipcode$white / zipcode$total_pop

race <- zipcode %>%
  select(ZIP_CODE, total_pop, white, black, ai, asian, hawaiian, other, two, p_black, p_white)

```

Get at zip code level for Oregon registered voters
```{r}
total_regs <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(count = n())
```

Proportion of registered that voted on Nov 2016 and Nov 2012 
```{r}
prop_voted2016 <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_v16 = mean(`11/08/2016` == "YES"))

prop_voted2012 <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_v12 = mean(`11/06/2012` == "YES"))
```
Proportion that are not Motor Voter Registered 
```{r}
a <- (is.na(voter_or$DESCRIPTION))
voter_or$DESCRIPTION[a] <- "conventional"
voter_or$DESCRIPTION <- as.factor(voter_or$DESCRIPTION)

table(voter_or$DESCRIPTION)

prop_regcon <- voter_or %>%
  group_by(ZIP_CODE) %>%
  summarize(prop_reg = mean(DESCRIPTION == "conventional"))

#Dont need this any more, also does not work. 
#votor_or2 <- votor_or %>%
#  mutate(DESCRIPTION = stringr::str_replace_all(DESCRIPTION, NA, "Not Registered")
         
#MV <- voter_or2 %>%
  #group_by(ZIP_CODE) %>%
  #summarize(prop_MV = mean(DESCRIPTION != "conventional"))
  
  
      # prop_motorvoter <- voter_or %>%
 # group_by(ZIP_CODE) %>%
 # summarize(prop_NotMV = mean(`DESCRIPTION` == "Not Registered"))
```

Construct data sets with Voter info by Zip Code including, number of registered voters, how they registered, and the proportion that voted

```{r}
zipcode_data <- inner_join(x = total_regs,
                           y = prop_voted2016,
                           by = "ZIP_CODE")

zipcode_data <- inner_join(x = zipcode_data,
                           y = prop_voted2012,
                           by = "ZIP_CODE")

zipcode_data <- inner_join(x = zipcode_data,
                           y = prop_regcon,
                           by = "ZIP_CODE")

# Join sex with voter reg aggregated data
sex_reg <- inner_join(x = sex, y = zipcode_data, 
                      by = "ZIP_CODE")

# Join race with voter reg aggregated data
race_reg <- inner_join(x = race, y = zipcode_data, 
                      by = "ZIP_CODE")
# One set with both 

county_reg <- inner_join(x = race_reg, y = sex, 
                      by = "ZIP_CODE")


```

Analysis 
```{r}
#I would like to have a percent black, and a percent white ect columns. Then the analysis I can do is- correlation between regisration through MV and percentage white. Percent vote and race.

```
Extra Code 
```{r}
#race_reg2<- race_reg %>% 
  #gather(key="race", value = "number", 3:9)

#sex_reg2 <- sex_reg %>% 
 # gather(key="sex", value = "number", 3:4)

#both_regtidy <-inner_join(x = race_reg2, y = sex_reg2, 
                    #  by = "ZIP_CODE") 


```

